---
title: "Storyboard"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---


```{r}
library(tidyverse)
source("compmus.R")

compmus2025 <- read_csv("compmus2025.csv")
```

```{r}
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

#major_key <-
#  c(5, 2, 3.5, 2, 4.5, 4.0, 2.0, 4.5, 2.0, 3.5, 1.5, 4.0)
#minor_key <-
#  c(5, 2, 3.5, 4.5, 2.0, 4.0, 2.0, 4.5, 3.5, 2.0, 1.5, 4.0)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )
```


Track 1 {data-navmenu="My Tracks" data-icon="ion-ios-musical-note"}
====================================================

Column {.tabset}
--------------------------------------------------------------------------


### Chromagram Track 1 

```{r}
"features/tobias-p-1.json" |>                           
  compmus_chroma(norm = "euclidean") |>                 
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()   
```


### Cepstrogram Track 1

```{r}
"features/tobias-p-1.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()
```


### Chroma-based self-similarity track 1

```{r}
"features/tobias-p-1.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "angular"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```


### Timbre-based self-similarity track 1

```{r}
"features/tobias-p-1.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "angular"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```



### Chordogram Track 1

```{r}
"features/tobias-p-1.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "euclidean",       # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "angular"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic() 
```

### keyagram Track 1

```{r}
"features/tobias-p-1.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    key_templates,         # Change to chord_templates if desired
    norm = "euclidean",       # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "angular"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic() 
```

Track 2 {data-navmenu="My Tracks" data-icon="ion-ios-musical-note"}
====================================================

Column {.tabset}
--------------------------------------------------------------------------

### Chromagram Track 2

```{r}
"features/tobias-p-2.json" |>                          
  compmus_chroma(norm = "euclidean") |>                
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +             
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()   
```

### Cepstrogram Track 2

```{r}
"features/tobias-p-2.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()
```


### Chroma-based self-similarity track 2

```{r}
"features/tobias-p-2.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "angular"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```


### Timbre-based self-similarity track 2

```{r}
"features/tobias-p-2.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "angular"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic() 
```


### Chordogram Track 2

```{r}
"features/tobias-p-2.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "euclidean",       # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "angular"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic() 
```

### keyagram Track 2

```{r}
"features/tobias-p-2.json" |> 
  compmus_chroma(norm = "euclidean") |> 
  compmus_match_pitch_templates(
    key_templates,         # Change to chord_templates if desired
    norm = "euclidean",       # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "angular"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic() 
```



Introduction
====================================================

In this portfolio we will be using a student musical corpus consisting of tracks submited by students of the course Computational Musicology.

```{r}
```

Visualisation
====================================================


Column {.tabset}
--------------------------------------------------------------------------

### Arousal

```{r}
ggplot(compmus2025, aes(x=tempo, y=arousal)) + geom_point()
```

### Danceability

```{r}
ggplot(compmus2025, aes(x=tempo, y=danceability)) + geom_point()
```

### Engagingness

```{r}
ggplot(compmus2025, aes(x=tempo, y=engagingness)) + geom_point()
```

### Instrumentalness

```{r}
ggplot(compmus2025, aes(x=tempo, y=instrumentalness)) + geom_point()
```

### Valence

```{r}
ggplot(compmus2025, aes(x=tempo, y=valence)) + geom_point()
```

Column {.tabset}
------------------------------------------------------
### Arousal x Engagingness

```{r}
ggplot(compmus2025, aes(x=arousal, y=engagingness)) + geom_point()
```

### Arousal x Instrumentalness

```{r}
ggplot(compmus2025, aes(x=arousal, y=instrumentalness)) + geom_point()
```

### Arousal x Danceability

```{r}
ggplot(compmus2025, aes(x=arousal, y=danceability)) + geom_point()
```



Conclusion
====================================================

```{r}
```
